---
title: "SafeALZ, t-test results"
author: "The SafeALZ Group"
output: html_document
---
 

```{r setup, include=FALSE} 

# Load packages
library(pacman)
p_load(knitr, dplyr, ggplot2, ggpubr, janitor, flextable, officer)

# Load the dataset
qtc_valmis <- read.csv("../data/manual_bazetts_qtcs.csv")


# Set up necessary general settings
knitr::opts_chunk$set(echo = TRUE)

```
 
 

<br>
<br>

# Comparison: is there a difference between QTc time before and after the initiation of ChE inhibitor treatment?

<br>

We can use the paired t-test to solve this question.



### SafeALZ: prepare data set (memantine not allowed)

<br>

Let's exclude all memantine users and make a new data set.

<br>

```{r prepare_safealz_dataset}
 
# create a new data set, exclude memantine users
article1 <- qtc_valmis %>% filter(!mlaake.factor == "memantiini")

# Filter out NA values

article1 <- article1 %>%  filter(!is.na(pre_ekg_qtcaika))
article1 <- article1 %>%  filter(!is.na(post_ekg_qtcaika))

# confirm no memantine users exist
prop.table(table(article1$mlaake.factor)) * 100

# tidy up the data set
article1 <- article1 %>% select(record_id, mlaake.factor, pre_ekg_qtcaika, post_ekg_qtcaika)

# shorten a bit the variable names

article1 <- article1 %>% rename(drug_used = mlaake.factor)
article1 <- article1 %>% rename(qtc_before = pre_ekg_qtcaika)
article1 <- article1 %>% rename(qtc_after = post_ekg_qtcaika)

head(article1)


article1$drug_used <- as.factor(article1$drug_used)

levels(article1$drug_used)



# Change labels of factor variable

levels(article1$drug_used)[levels(article1$drug_used) == "donepetsiili"] <- "donepezil"
levels(article1$drug_used)[levels(article1$drug_used) == "galantamiini"] <- "galantamine"
levels(article1$drug_used)[levels(article1$drug_used) == "rivastigmiini_tbl"] <- "rivastigmine tablet"
levels(article1$drug_used)[levels(article1$drug_used) == "rivastigmiini_laast"] <- "rivastigmine skin patch"
levels(article1$drug_used)[levels(article1$drug_used) == "memantiini"] <- "memantine"


# Confirm that the levels are correct

levels(article1$drug_used)


# Store the data to new data frame

qtc_data   <-    article1



```


<br>
<br>


### Introduce custom functions to round statistics for tables

<br>

In order to have the exact desired decimal places in the tables, it is worthwhile to use custom functions which enable to round the statistics in a custom way for the tables.

(The code for that is in a hidden code chunk below this line.)

<br>

```{r stylize_statistics, echo=FALSE}
 
# create a custom function that stylizes the statistic values

stylize2 <- function(x) {
  case_when(
    x < -0.01 ~ as.character(format(round_half_up(x, 2), nsmall = 2)),
    
    x >= -0.01 & x <= -0.001 & round_half_up(x, 2) == "-0.01"
    ~ as.character(format(round_half_up(x, 2), nsmall = 2)),

    x >= -0.01 & x <= -0.001 & round_half_up(x, 2) != "-0.01"
    ~ as.character(format(round_half_up(x, 3), nsmall = 3)),

    x > -0.001 & x < 0 ~ "-<.001",
    x == 0 ~ "0.00",
    x >= 0 & x < 0.001 ~ "<.001",

    x >= 0.001 & x < 0.005 ~ as.character(format(round_half_up(x, 3), nsmall = 3)),

    x >= 0.005 ~ as.character(format(round_half_up(x, 2), nsmall = 2)),
    
    TRUE ~ "other"
  )
}


stylize1 <- function(x) {
  case_when(
    x <= -0.05 ~ as.character(format(round_half_up(x, 1), nsmall = 1)),
    x < -0.01 & x > -0.05 ~ as.character(format(round_half_up(x, 2), nsmall = 1)),
    x >= -0.01 & x <= -0.001 & round_half_up(x, 2) == "-0.01"
    ~ as.character(format(round_half_up(x, 2), nsmall = 2)),

    x >= -0.01 & x <= -0.001 & round_half_up(x, 2) != "-0.01"
    ~ as.character(format(round_half_up(x, 3), nsmall = 3)),

    x > -0.001 & x < 0 ~ "-<.001",
    x == 0 ~ "0.00",
    x == -0 ~ "0.00",
    x >= 0 & x < 0.001 ~ "<.001",

    x >= 0.001 & x < 0.005 ~ as.character(format(round_half_up(x, 3), nsmall = 3)),

    x >= 0.005 & x < 0.05 ~ as.character(format(round_half_up(x, 2), nsmall = 2)),
    x >= 0.05 ~ as.character(format(round_half_up(x, 1), nsmall = 1)),
    
    TRUE ~ "other"
  )
}



# create a custom function that stylizes p-values
#
# (p-values shall not land at 0.05 sharp)

stylize_p <- function(x) {
  case_when(
    x < 0.001 ~ "<.001",
    x >= 0.001 & x < 0.005 ~ as.character(format(round_half_up(x, 3), nsmall = 3)),

    x >= 0.005 & x < 0.045 ~ as.character(format(round_half_up(x, 2), nsmall = 2)),
    x >= 0.045 & x < 0.055 ~ as.character(format(round_half_up(x, 3), nsmall = 3)),
    x >= 0.055 ~ as.character(format(round_half_up(x, 2), nsmall = 2)),
    TRUE ~ "other"
  )
}


```


<br>
<br>


<!-- ### Let's produce some random data for drugs and QTc -->

<!-- <br> -->


<!-- ```{r random_data} -->


<!-- # Set the number of participants -->
<!-- n_participants <- 50 -->

<!-- # Create a data frame with participant IDs -->
<!-- qtc_data <- data.frame(participant_id = 1:n_participants) -->

<!-- # Define the drugs -->
<!-- drugs <- c("donepezil", "memantine", "galantamine", "rivastigmine tablet", "rivastigmine skin patch") -->

<!-- # Generate random data for the drugs used and QTc measurements -->
<!-- qtc_data$drug_used <- sample(drugs, n_participants, replace = TRUE) -->
<!-- qtc_data$drug_used <- as.factor(qtc_data$drug_used) -->

<!-- qtc_data$qtc_before <- rnorm(n_participants, mean = 400, sd = 20) -->
<!-- qtc_data$qtc_after <- rnorm(n_participants, mean = 450, sd = 30) -->

<!-- # View the resulting data frame -->
<!-- head(qtc_data) -->

<!-- ``` -->

<!-- <br> -->
<!-- <br> -->




### Exclusion of memantine users

<br>

Let's exclude memantine users.

<br>

```{r exclude_memantine}
 

qtc_data <- qtc_data %>% filter(!drug_used == "memantine")

# confirm no memantine users exist

prop.table(table(qtc_data$drug_used)) * 100


```


<br>
<br>
 
 
 


### T-test results - all ChEIs

<br>

```{r all_AChEIs_t_test}
 

# run t-test
t_test_results <- t.test(qtc_data$qtc_before, qtc_data$qtc_after, paired=TRUE)

# extract p-value
p_value <- t_test_results$p.value

# extract mean difference
mean_diff <- t_test_results$estimate[1]

# extract t-value
t_value <- t_test_results$statistic["t"]


# print t-test results
t_test_results


# define n (number of participants)

n <- qtc_data %>% nrow()

# calculate mean for qtc_before
qtcbefore_mean <- mean(qtc_data$qtc_before)


# calculate SD for qtc_before
qtcbefore_sd <- sd(qtc_data$qtc_before)


# calculate mean for qtc_after
qtcafter_mean <- mean(qtc_data$qtc_after)

#calculate SD for qtc_after
qtcafter_sd <- sd(qtc_data$qtc_after)


#create a tibble of the results

All <- tibble::tribble(
  
  ~Group,       ~n, ~qtcbefore_mean, ~qtcbefore_sd, ~qtcafter_mean, ~qtcafter_sd, ~mean_diff, ~t_value,     ~p_value,
  
  "All ChEIs",  n,  qtcbefore_mean,  qtcbefore_sd,  qtcafter_mean,  qtcafter_sd,   mean_diff,         t_value,   p_value)


All %>% kable()

# destroy the statistics so that they can be run with confidence
# for the next group in the next chunk

rm(t_test_results, n, qtcbefore_mean, qtcbefore_sd, qtcafter_mean, qtcafter_sd, mean_diff, t_value, p_value)



```


<br>
<br>



### T-test results - donepezil

<br>


```{r donepezil_t_test}

# select only donepezil users
done <- qtc_data %>% filter(drug_used == "donepezil")

# confirm that only donepezil users exist

prop.table(table(done$drug_used)) * 100


# run t-test
t_test_results <- t.test(done$qtc_before, done$qtc_after, paired=TRUE)

# extract p-value
p_value <- t_test_results$p.value

# extract mean difference
mean_diff <- t_test_results$estimate[1]

# extract t-value
t_value <- t_test_results$statistic["t"]

# print t-test results
t_test_results


# define n (number of participants)

n <- done %>% nrow()

# calculate mean for qtc_before
qtcbefore_mean <- mean(done$qtc_before)

# calculate SD for qtc_before
qtcbefore_sd <- sd(done$qtc_before)


# calculate mean for qtc_after
qtcafter_mean <- mean(done$qtc_after)

#calculate SD for qtc_after
qtcafter_sd <- sd(done$qtc_after)




#create a tibble of the results

Donepezil <- tibble::tribble(
  
  ~Group,       ~n, ~qtcbefore_mean, ~qtcbefore_sd, ~qtcafter_mean, ~qtcafter_sd, ~mean_diff, ~t_value,     ~p_value,
  "Donepezil",  n,  qtcbefore_mean,  qtcbefore_sd,  qtcafter_mean,  qtcafter_sd,   mean_diff,         t_value,   p_value)


Donepezil %>% kable()



# destroy the statistics so that they can be run with confidence
# for the next group in the next chunk

rm(t_test_results, n, qtcbefore_mean, qtcbefore_sd, qtcafter_mean, qtcafter_sd, mean_diff, t_value, p_value)



```

 

<br>
<br>



### T-test results - galantamine

<br>




```{r galantamine_t_test}

# select only galantamine users
gala <- qtc_data %>% filter(drug_used == "galantamine")

# confirm that only galantamine users exist

prop.table(table(gala$drug_used)) * 100


# run t-test
t_test_results <- t.test(gala$qtc_before, gala$qtc_after, paired=TRUE)

# extract p-value
p_value <- t_test_results$p.value

# extract mean difference
mean_diff <- t_test_results$estimate[1]

# extract t-value
t_value <- t_test_results$statistic["t"]

# print t-test results
t_test_results


# define n (number of participants)

n <- gala %>% nrow()

# calculate mean for qtc_before
qtcbefore_mean <- mean(gala$qtc_before)

# calculate SD for qtc_before
qtcbefore_sd <- sd(gala$qtc_before)


# calculate mean for qtc_after
qtcafter_mean <- mean(gala$qtc_after)

#calculate SD for qtc_after
qtcafter_sd <- sd(gala$qtc_after)

 


#create a tibble of the results

Galantamine <- tibble::tribble(
  
  ~Group,       ~n, ~qtcbefore_mean, ~qtcbefore_sd, ~qtcafter_mean, ~qtcafter_sd, ~mean_diff, ~t_value,     ~p_value,
  "Galantamine",  n,  qtcbefore_mean,  qtcbefore_sd,  qtcafter_mean,  qtcafter_sd,   mean_diff,         t_value,   p_value)


Galantamine %>% kable()



# destroy the statistics so that they can be run with confidence
# for the next group in the next chunk

rm(t_test_results, n, qtcbefore_mean, qtcbefore_sd, qtcafter_mean, qtcafter_sd, mean_diff, t_value, p_value)



```




<br>
<br>



### T-test results - rivastigmine

<br>



```{r rivastigmine_t_test}

# select only rivastigmine users

riva <- qtc_data %>% filter(drug_used == "rivastigmine tablet" | drug_used == "rivastigmine skin patch")

# confirm that only rivastigmine users exist

prop.table(table(riva$drug_used)) * 100





# run t-test
t_test_results <- t.test(riva$qtc_before, riva$qtc_after, paired=TRUE)

# extract p-value
p_value <- t_test_results$p.value

# extract mean difference
mean_diff <- t_test_results$estimate[1]

# extract t-value
t_value <- t_test_results$statistic["t"]

# print t-test results
t_test_results


# define n (number of participants)

n <- riva %>% nrow()

# calculate mean for qtc_before
qtcbefore_mean <- mean(riva$qtc_before)

# calculate SD for qtc_before
qtcbefore_sd <- sd(riva$qtc_before)


# calculate mean for qtc_after
qtcafter_mean <- mean(riva$qtc_after)

#calculate SD for qtc_after
qtcafter_sd <- sd(riva$qtc_after)

 


#create a tibble of the results

Rivastigmine <- tibble::tribble(
  
  ~Group,       ~n, ~qtcbefore_mean, ~qtcbefore_sd, ~qtcafter_mean, ~qtcafter_sd, ~mean_diff, ~t_value,     ~p_value,
  "Rivastigmine",  n,  qtcbefore_mean,  qtcbefore_sd,  qtcafter_mean,  qtcafter_sd,   mean_diff,         t_value,   p_value)


Rivastigmine %>% kable()



# destroy the statistics so that they can be run with confidence
# for the next group in the next chunk

rm(t_test_results, n, qtcbefore_mean, qtcbefore_sd, qtcafter_mean, qtcafter_sd, mean_diff, t_value, p_value)



```




<br>
<br>



### Combine the statistics results for all groups: All, donepezil, galantamine, rivastigmine

<br>



```{r combine_results}

results <- rbind(All, Donepezil, Galantamine, Rivastigmine)

results

results %>% kable()


```



<br>
<br>



### Stylize (round) the statistics

<br>



```{r stylize_results}

# Apply a function to specific columns of the results

results <- results %>% mutate(across
                          (qtcbefore_mean:mean_diff, stylize1))

results <- results %>% mutate(t_value = stylize2(t_value))

results <- results %>% mutate(p_value = stylize_p(p_value))

results <- results %>% mutate(n = as.character(n))
 

# Verify how results look

results

results %>% kable()

 
```



<br>
<br>



### Give the statistics nicer names

<br>



```{r nicer_names_to_variables}

 
results <- results %>% rename(`Mean difference` =  mean_diff,
                              `t` =  t_value,
                              `p-value` = p_value)
    

```



<br>
<br>



### Create a simple flextable of the results

<br>



```{r simple_qtc_data_flextable}

results %>% flextable()





```



<br>
<br>



### Create a more advanced flextable of the results

<br>



```{r advanced_qtc_data_flextable_take2}

 
results2 <- results


# Add empty row

emptyrow <- tibble::tribble(
  
  ~Group,       ~n, ~qtcbefore_mean, ~qtcbefore_sd, ~qtcafter_mean, ~qtcafter_sd, ~`Mean difference`, ~`t`,     ~`p-value`,
  "Overall",  "","","","","","","","")

 
# Reorder rows
results2 <- bind_rows(slice(results2, 2:n()), emptyrow, slice(results2, 1))


# Create empty column

` ` <- c("","","","","")

` ` <- tibble(` `)
 

# Reorder columns
results2 <- bind_cols(select(results2, 1:4), ` `, select(results2, 5:9))

# Turn the data set into a flextable object
ft <- flextable(results2)

# First delete the header
ft <- delete_part(x = ft, part = "header")

# Create new header, 1st row
ft <- add_header_row(
  ft,
  values = c("Group", "n", "Mean", "SD", "", "Mean", "SD", "Mean difference", "t", "p-value"),
  top = FALSE
)

# Add a 2nd row of the multilevel header
ft <- add_header_row(ft,
                     values = c("", "QTc before medication", "", "QTc during medication", ""),
                     colwidths = c(2, 2, 1, 2, 3), top = TRUE)

# # Add a 3rd row of the multilevel header
# ft <- add_header_row(ft, values = c("", "QTc interval", ""),
#                      colwidths = c(2, 5, 3), top = TRUE)
# 
# ft


# Add indentation to all rows except to "Overall"
ft <- ft %>% prepend_chunks(i = ~ Group != "Overall", j = "Group", as_chunk("\t"))


# Add theme (this cannot be done later as it messes up with the align commands later on)
ft <- theme_booktabs(ft)

# Align first 2 columns to left
ft <- align(ft, j = 1, align = "left", part = "all")

# Align the rest to center
ft <- align(ft, j = 2:10, align = "center", part = "all")

# Add custom borders under multi-level headers
#
# First define define the style of the border line
border_style = officer::fp_border(color="black", width=1.5)

ft <- hline(ft, part="header", i = 1, j = 3, border = border_style)
#ft <- hline(ft, part="header", i = 2, j = 3, border = border_style)
ft <- hline(ft, part="header", i = 1, j = 6, border = border_style)


#Squeeze (autofit) the flextable
ft <- ft %>% autofit()

# Modify column widths
ft <- width(ft, j=c(3:4, 6:7), width = 1)

# Modify more
ft <- width(ft, j = ~ ` `, width = 0.001)

# Stylize the p-value to italic
ft <- mk_par(ft,
    part = "header", j = "p-value",
    value = as_paragraph(as_i("p"), "-value")
  )

# Stylize the t to italic
ft <- mk_par(ft,
    part = "header", j = "t",
    value = as_paragraph(as_i("t"))
  )

# Stylize the n to italic
ft <- mk_par(ft,
    part = "header", j = "n",
    value = as_paragraph(as_i("n"))
  )



#Change the header to bold
ft <- ft %>%
  bold(part = "header", bold = TRUE)

#Change the subheaders in column 1 to bold
ft <- ft %>%
  bold(j = 1,
       i = 4,
       part = "body",
       bold = TRUE)

#Stylize 1st SD to italic
ft <- ft %>%
  italic(j = 4,
       i = 2,
       part = "header",
       italic = TRUE)

#Stylize 2nd SD to italic
ft <- ft %>%
  italic(j = 7,
       i = 2,
       part = "header",
       italic = TRUE)

# Add a header
ft <- ft %>% add_header_lines("Table 2. QTc values measured before and during ChEI medication")


# Add a footer
ft <- ft %>% add_footer_lines(c(

  "QTc, corrected QT interval; ChEI, cholinesterase inhibitor."))


# Print the flextable
ft

# Define the .docx qualities

sect_properties <- prop_section(
  page_size = page_size(orient = "landscape")
)

# Save the flextable to docx

ft %>% save_as_docx(path = "table2_csv.docx", pr_section = sect_properties)

```



<br>
<br>

