---
title: "SafeALZ, t-test results"
author: "The SafeALZ Group"
output: html_document
---
 

```{r setup, include=FALSE} 

# Load packages
library(pacman)
p_load(knitr, dplyr, ggplot2, ggpubr, gridExtra, cowplot, patchwork, janitor, grid)
p_load(ggthemes, showtext)
p_load(animation, magick, pdftools)

# Load the dataset
qtc_valmis <- read.csv("../data/manual_bazetts_qtcs.csv")

# Set up necessary general settings

knitr::opts_chunk$set(echo = TRUE)

# # Load specific font from Google Fonts
# font_add_google("Rosario", "google")

# Load specific font from Google Fonts
font_add_google("Lato", "google")

# Invoke showtext
showtext_auto()


# Name colors

mycolor1 = "#F2D7B2" # "Wheat"
mycolor2 = "#CFBCD8" # "Lavender grey"

```
 
 

<br>
<br>

# Comparison: is there a difference between QTc time before and after the initiation of ChE inhibitor treatment?

<br>

We can use the paired t-test to solve this question.




### SafeALZ: prepare data for graph

<br>

We will now:

1. Choose only the necessary variables for the graphs.

2. Exclude participants with missing data for QTc interval time.

3. "Stack" the data into a "long" format

4. Give variables names that are displayed in the ready graphs.

<br>
<br>

```{r safealz_prepare_data}
 

# Copy the data to a new data frame, select only some variables

qtc_valmis2 <- qtc_valmis %>% select(record_id, mlaake, mlaake.factor, pre_ekg_qtcaika, post_ekg_qtcaika) 

# Filter out NA values

qtc_valmis2 <- qtc_valmis2 %>%  filter(!is.na(pre_ekg_qtcaika))
qtc_valmis2 <- qtc_valmis2 %>%  filter(!is.na(post_ekg_qtcaika))

# Give variables nicer names

qtc_valmis2 <- qtc_valmis2 %>%  rename(`Before medication` = pre_ekg_qtcaika)
qtc_valmis2 <- qtc_valmis2 %>%  rename(`During medication` = post_ekg_qtcaika)

# `Before medication` and `During medication` have a slightly curious
# variable type "labelled integer" which can't be stacked.
# Fix that first.

qtc_valmis2 <- qtc_valmis2 %>%
  mutate(`Before medication` = as.numeric(`Before medication`),
         `During medication` = as.numeric(`During medication`))



```

<br>
<br>
<br>
<br>
<br>
<br>



### Introduce a custom function to round p-value

<br>

(The code for that is in a hidden code chunk below this line.)

<br>

```{r stylize_statistics, echo=FALSE}
 

# create a custom function that stylizes p-values
#
# (p-values shall not land at 0.05 sharp)

stylize_p <- function(x) {
  case_when(
    x < 0.001 ~ "<.001",
    x >= 0.001 & x < 0.005 ~ as.character(format(round_half_up(x, 3), nsmall = 3)),

    x >= 0.005 & x < 0.045 ~ as.character(format(round_half_up(x, 2), nsmall = 2)),
    x >= 0.045 & x < 0.055 ~ as.character(format(round_half_up(x, 3), nsmall = 3)),
    x >= 0.055 ~ as.character(format(round_half_up(x, 2), nsmall = 2)),
    TRUE ~ "other"
  )
}


```





### SafeALZ real data: boxplot + t-test results (all ChEI users)

 
<br>
<br>
<br>
<br>


```{r all_achei_users_graph, echo=FALSE}



#################################################################
##           Filter out users of memantine                     ##
#################################################################


all_acheis <- qtc_valmis2 %>% filter(!mlaake.factor == "memantiini")

n_all <- all_acheis %>% nrow()

n_all

# If wanted, a probability table can be shown to test that users of all
# medications exist in the data set

prop.table(table(all_acheis$mlaake.factor)) * 100



#################################################################
##            Give the plot a title below this text            ##
#################################################################

title <- paste0("All ChEI users (n=", n_all, ")")






#################################################################
##                Stack the data to long format                ##
#################################################################


# Stack the data to a "long" format
#
# (The numbers represent the variables that are chosen for this)

all_acheis <- as.data.frame(all_acheis)

stacked_data <- stack(all_acheis[,c("Before medication","During medication")])



# If wanted, the head and tail of the long form data can be shown

#head(stacked_data)
#tail(stacked_data)

  
###########################################################################
###########################################################################
###                                                                     ###
###               BELOW THIS TEXT: THE CODE FOR THE GRAPH               ###
###                                                                     ###
###########################################################################
###########################################################################




# Create the "canvas" for the ggplot; save it to an object called "bp"

bp <- ggplot(stacked_data, aes(x=ind, y=values, fill=ind)) +
  
  #ylim(350, 550) +
  #scale_y_continuous(limits=c(375, 525), breaks = seq(400, 500, by = 25), expand=c(0, 0)) +
  scale_y_continuous(limits=c(350, 550), expand=c(0,0)) +

  ggtitle(title) +                                  # Add a title to the plot
  geom_boxplot() +                                  # Draw the boxes
  theme_classic(base_size = 16, base_family = "google") +  # Classic theme
  theme(legend.position = "none") +                 # Remove the legend
  labs(x = "Time point of ECG") +                   # Change the X-axis title
  labs(y = "QTc interval (ms)") +                   # Change the Y-axis title
  theme(plot.title = element_text(size = 16, face = "bold")) +     # Title font size
  theme(axis.text.x = element_text(size = 14)) +    # Increase X-tick label size
  theme(axis.text.y = element_text(size = 14)) +    # Increase X-tick label size

  theme(axis.title.x = element_text(size = 16,                 # Modify X-axis label
                                    margin = margin(t = 20),
                                    face = "bold")) +

  theme(axis.title.y = element_text(size = 16,                 # Modify Y-axis label
                                    margin = margin(r = 15),
                                    face = "bold")) +

  theme(axis.text.x = element_text(margin = margin(t = 10))) +  # Set distance for the X-axis tick labels
  theme(axis.text.y = element_text(margin = margin(r = 5))) +   # Set distance for the Y-axis tick labels

  theme(plot.margin = unit(c(1,1,1,1), "cm")) + # Modify plot margin

   scale_fill_manual(values=c(mycolor1, mycolor2))   # Manual fill colors


# Add boxplot to the object "bp"; save the result to an object called "p4"

p4 <- bp + stat_compare_means(
  method = "t.test",
  label = "p.format",
  size = 5,
  show.legend = FALSE,
  hide.ns = FALSE,
  paired = TRUE,
  comparisons = list(c("Before medication", "During medication")) ,
  method.args = list(symbol = TRUE)
)


# Change the font of the comparison
p4$layers[[2]]$aes_params$family <- "google"


# Call the object "p4" (essentially: display the graph)
#p4


## Extract data with ggplot_build
p4build <- ggplot_build(p4)

## Explore build data to find the right variable (here commented out)
#p4build$data

p_p4 <- p4build$data[[2]]$annotation %>% as.numeric() %>% stylize_p


# Replace the p-value with the rounded one
p4build$data[[2]]$annotation <- paste0("P = ", p_p4)

## plot anew
plot(ggplot_gtable(p4build))


p4 <- ggplot_gtable(p4build)

p4 <- cowplot::ggdraw(p4)



```

<br>
<br>
<br>
<br>
<br>
<br>

### SafeALZ real data: boxplot + t-test results (donepezil users)

<br>
<br>
<br>
<br>

```{r donepezil_users_graph, echo=FALSE}



#################################################################
##           Filter only users of certain medication           ##
#################################################################


done <- qtc_valmis2 %>% filter(mlaake.factor == "donepetsiili")

n_done <- qtc_valmis2 %>% filter(mlaake.factor == "donepetsiili") %>% nrow()

# If wanted, a probability table can be shown to test that only
# donepezil users remain

prop.table(table(done$mlaake.factor)) * 100

#################################################################
##            Give the plot a title below this text            ##
#################################################################

title <- paste0("Donepezil users (n=", n_done, ")")



#################################################################
##                Stack the data to long format                ##
#################################################################


# Stack the data to a "long" format
#
# (The numbers represent the variables that are chosen for this)

done <- as.data.frame(done)

stacked_data <- stack(done[,c("Before medication","During medication")])



# If wanted, the head and tail of the long form data can be shown

#head(stacked_data)
#tail(stacked_data)


###########################################################################
###########################################################################
###                                                                     ###
###               BELOW THIS TEXT: THE CODE FOR THE GRAPH               ###
###                                                                     ###
###########################################################################
###########################################################################
 



# Create the "canvas" for the ggplot; save it to an object called "bp"

bp <- ggplot(stacked_data, aes(x=ind, y=values, fill=ind)) +
  
  #ylim(350, 550) +
  #scale_y_continuous(breaks = seq(400, 500, by = 50)) +
  scale_y_continuous(limits=c(350, 550), expand=c(0,0)) +

  ggtitle(title) +                                  # Add a title to the plot
  geom_boxplot() +                                  # Draw the boxes
  theme_classic(base_size = 16, base_family = "google") +  # Classic theme
  theme(legend.position = "none") +                 # Remove the legend
  labs(x = "Time point of ECG") +                   # Change the X-axis title
  labs(y = "QTc interval (ms)") +                   # Change the Y-axis title
  theme(plot.title = element_text(size = 16, face = "bold")) +     # Title font size
  theme(axis.text.x = element_text(size = 14)) +    # Increase X-tick label size
  theme(axis.text.y = element_text(size = 14)) +    # Increase X-tick label size

  theme(axis.title.x = element_text(size = 16,                 # Modify X-axis label
                                    margin = margin(t = 20),
                                    face = "bold")) +

  theme(axis.title.y = element_text(size = 16,                 # Modify Y-axis label
                                    margin = margin(r = 15),
                                    face = "bold")) +

  theme(axis.text.x = element_text(margin = margin(t = 10))) +  # Set distance for the X-axis tick labels
  theme(axis.text.y = element_text(margin = margin(r = 5))) +   # Set distance for the Y-axis tick labels

  theme(plot.margin = unit(c(1,1,1,1), "cm")) + # Modify plot margin

   scale_fill_manual(values=c(mycolor1, mycolor2))   # Manual fill colors
  
# Add boxplot to the object "bp"; save the result to an object called "p1"

p1 <- bp + stat_compare_means(
  method = "t.test",
  label = "p.format",
  size = 5,
  show.legend = FALSE,
  hide.ns = FALSE,
  paired = TRUE,
  comparisons = list(c("Before medication", "During medication")),
  method.args = list(symbol = TRUE)
)


# Change the font of the comparison
p1$layers[[2]]$aes_params$family <- "google"


# Call the object "p1" (essentially: display the graph)

#p1
#
#
#
## Extract data with ggplot_build
 p1build <- ggplot_build(p1)
#
## Explore build data to find the right variable (here commented out)
#p1build$data

p_p1 <- p1build$data[[2]]$annotation %>% as.numeric() %>% stylize_p
#
#
# # Replace the p-value with the rounded one
p1build$data[[2]]$annotation <- paste0("P = ", p_p1)
#
## plot anew
plot(ggplot_gtable(p1build))
#
#
#
p1 <- ggplot_gtable(p1build)
#
p1 <- cowplot::ggdraw(p1)
#


```

<br>
<br>
<br>
<br>
<br>
<br>

### SafeALZ real data: boxplot + t-test results (galantamine users)

<br>
<br>
<br>
<br>

```{r galantamine_users_graph, echo=FALSE}



#################################################################
##           Filter only users of certain medication           ##
#################################################################


gala <- qtc_valmis2 %>% filter(mlaake.factor == "galantamiini")

n_gala <- qtc_valmis2 %>% filter(mlaake.factor == "galantamiini") %>% nrow()

# If wanted, a probability table can be shown to test that only
# donepezil users remain

prop.table(table(gala$mlaake.factor)) * 100

#################################################################
##            Give the plot a title below this text            ##
#################################################################

title <- paste0("Galantamine users (n=", n_gala, ")")



#################################################################
##                Stack the data to long format                ##
#################################################################


# Stack the data to a "long" format
#
# (The numbers represent the variables that are chosen for this)

gala <- as.data.frame(gala)

stacked_data <- stack(gala[,c("Before medication","During medication")])



# If wanted, the head and tail of the long form data can be shown

#head(stacked_data)
#tail(stacked_data)


###########################################################################
###########################################################################
###                                                                     ###
###               BELOW THIS TEXT: THE CODE FOR THE GRAPH               ###
###                                                                     ###
###########################################################################
###########################################################################

 


# Create the "canvas" for the ggplot; save it to an object called "bp"

bp <- ggplot(stacked_data, aes(x=ind, y=values, fill=ind)) +

  #ylim(350, 550) +
  #scale_y_continuous(breaks = seq(400, 500, by = 50)) +
  scale_y_continuous(limits=c(350, 550), expand=c(0,0)) +

  ggtitle(title) +                                  # Add a title to the plot
  geom_boxplot() +                                  # Draw the boxes
  theme_classic(base_size = 16, base_family = "google") +  # Classic theme
  theme(legend.position = "none") +                 # Remove the legend
  labs(x = "Time point of ECG") +                   # Change the X-axis title
  labs(y = "QTc interval (ms)") +                   # Change the Y-axis title
  theme(plot.title = element_text(size = 16, face = "bold")) +     # Title font size
  theme(axis.text.x = element_text(size = 14)) +    # Increase X-tick label size
  theme(axis.text.y = element_text(size = 14)) +    # Increase X-tick label size

  theme(axis.title.x = element_text(size = 16,                 # Modify X-axis label
                                    margin = margin(t = 20),
                                    face = "bold")) +

  theme(axis.title.y = element_text(size = 16,                 # Modify Y-axis label
                                    margin = margin(r = 15),
                                    face = "bold")) +

  theme(axis.text.x = element_text(margin = margin(t = 10))) +  # Set distance for the X-axis tick labels
  theme(axis.text.y = element_text(margin = margin(r = 5))) +   # Set distance for the Y-axis tick labels

  theme(plot.margin = unit(c(1,1,1,1), "cm"))  + # Modify plot margin

 scale_fill_manual(values=c(mycolor1, mycolor2))   # Manual fill colors  
  
# Add boxplot to the object "bp"; save the result to an object called "p2"

p2 <- bp + stat_compare_means(
  method = "t.test",
  label = "p.format",
  size = 5,
  show.legend = FALSE,
  hide.ns = FALSE,
  paired = TRUE,
  comparisons = list(c("Before medication", "During medication")),
  method.args = list(symbol = TRUE)
)


# Change the font of the comparison
p2$layers[[2]]$aes_params$family <- "google"


# Call the object "p2" (essentially: display the graph)

#p2



## Extract data with ggplot_build
p2build <- ggplot_build(p2)

## Explore build data to find the right variable (here commented out)
#p2build$data

p_p2 <- p2build$data[[2]]$annotation %>% as.numeric() %>% stylize_p


# Replace the p-value with the rounded one
p2build$data[[2]]$annotation <- paste0("P = ", p_p2)

## plot anew
plot(ggplot_gtable(p2build))



p2 <- ggplot_gtable(p2build)

p2 <- cowplot::ggdraw(p2)



```

<br>
<br>
<br>
<br>
<br>
<br>



### SafeALZ real data: boxplot + t-test results (rivastigmine users)

<br>
<br>
<br>
<br>

```{r rivastigmine_users_graph, echo=FALSE}


#################################################################
##           Filter only users of certain medication           ##
#################################################################


riva <- qtc_valmis2 %>% filter(mlaake.factor == "rivastigmiini_tbl" | mlaake.factor == "rivastigmiini_laast")

n_riva <- qtc_valmis2 %>% filter(mlaake.factor == "rivastigmiini_tbl"
                                 | mlaake.factor == "rivastigmiini_laast") %>% nrow()

# If wanted, a probability table can be shown to test that only
# donepezil users remain

prop.table(table(riva$mlaake.factor)) * 100


#################################################################
##            Give the plot a title below this text            ##
#################################################################

title <- paste0("Rivastigmine users (n=", n_riva, ")")


#################################################################
##                Stack the data to long format                ##
#################################################################


# Stack the data to a "long" format
#
# (The numbers represent the variables that are chosen for this)

riva <- as.data.frame(riva)

stacked_data <- stack(riva[,c("Before medication","During medication")])



# If wanted, the head and tail of the long form data can be shown

head(stacked_data)
tail(stacked_data)


###########################################################################
###########################################################################
###                                                                     ###
###               BELOW THIS TEXT: THE CODE FOR THE GRAPH               ###
###                                                                     ###
###########################################################################
###########################################################################


# Create the "canvas" for the ggplot; save it to an object called "bp"

bp <- ggplot(stacked_data, aes(x=ind, y=values, fill=ind)) +

  #ylim(350, 550) +
  #scale_y_continuous(breaks = seq(400, 500, by = 50)) +
  #scale_y_continuous(limits=c(375, 525), breaks = seq(400, 500, by = 25), expand=c(0, 0)) +

  scale_y_continuous(limits=c(350, 550), expand=c(0,0)) +

  ggtitle(title) +                                  # Add a title to the plot
  geom_boxplot() +                                  # Draw the boxes
  theme_classic(base_size = 16, base_family = "google") +  # Classic theme
  theme(legend.position = "none") +                 # Remove the legend
  labs(x = "Time point of ECG") +                   # Change the X-axis title
  labs(y = "QTc interval (ms)") +                   # Change the Y-axis title
  theme(plot.title = element_text(size = 16, face = "bold")) +     # Title font size
  theme(axis.text.x = element_text(size = 14)) +    # Increase X-tick label size
  theme(axis.text.y = element_text(size = 14)) +    # Increase X-tick label size

  theme(axis.title.x = element_text(size = 16,                 # Modify X-axis label
                                    margin = margin(t = 20),
                                    face = "bold")) +

  theme(axis.title.y = element_text(size = 16,                 # Modify Y-axis label
                                    margin = margin(r = 15),
                                    face = "bold")) +

  theme(axis.text.x = element_text(margin = margin(t = 10))) +  # Set distance for the X-axis tick labels
  theme(axis.text.y = element_text(margin = margin(r = 5))) +   # Set distance for the Y-axis tick labels

  theme(plot.margin = unit(c(1,1,1,1), "cm")) +  # Modify plot margin

 scale_fill_manual(values=c(mycolor1, mycolor2))   # Manual fill colors



# Add boxplot to the object "bp"; save the result to an object called "p3"

p3 <- bp + stat_compare_means(
  method = "t.test",
  label = "p.format",
  size = 5,
  show.legend = FALSE,
  hide.ns = FALSE,
  paired = TRUE,
  comparisons = list(c("Before medication", "During medication")),
  method.args = list(symbol = TRUE)
)


# Call the object "p2" (essentially: display the graph)

#p3


# Change the font of the comparison
p3$layers[[2]]$aes_params$family <- "google"


## Extract data with ggplot_build
p3build <- ggplot_build(p3)

## Explore build data to find the right variable (here commented out)
#p3build$data

p_p3 <- p3build$data[[2]]$annotation %>% as.numeric() %>% stylize_p


# Replace the p-value with the rounded one
p3build$data[[2]]$annotation <- paste0("P = ", p_p3)

## plot anew
plot(ggplot_gtable(p3build))

p3 <- ggplot_gtable(p3build)

p3 <- cowplot::ggdraw(p3)


```

<br>
<br>
<br>
<br>
<br>
<br>



### A panel of all little graphs

<br>

Finally, we can present all little graphs in a panel.

The code below

1. First, creates the panel and saves as a PNG file
2. Then displays the PNG file

<br>


```{r panel, echo=FALSE}



# Create a panel

grid <- plot_grid(p1, p2, p3, p4, nrow = 2, ncol = 2, scale = 1.0)


# When showtext() is used to change font(s), plots have to be saved as pdf files,
# otherwise proportions get distorted.


# Save the panel to a PDF file
save_plot("fig1.pdf", grid, nrow = 3, ncol = 2)

# Load that pdf file with the magick package
panel <- magick::image_read_pdf("fig1.pdf")

# Save it as PNG
image_write(panel, path = "fig1.png", format = "png")


```


```{r show_panel, echo=FALSE}

# Display the ready panel
panel

```


